---
title: "Data Management"
author: "Marissa Lee"
date: "September 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages
library(ggplot2)
library(grid)

```


### Why is data management important?

- Fosters reproducible science  
- Saves you from losing and/or duplicating things  
- Makes it easy to share when you need help or need to help others  
- Because you need to comply with funding agencies  

  
### An ideal workflow  

1. Have a meta-data file for each data file
2. Import raw data and never edit raw data files; think of them as precious artifacts
3. Make and use code scripts as much as possible
4. Annotate your code so that the human running the computer knows what is going on


### 1. More on meta-data

- Meta-data is just information used to describe your data, there are 3 types...
-**Descriptive** -- what common fields like column names mean, identify the datatypes in a column, identify the range of values in a column
-**Structural** -- how different pieces of data relate to one another, such as in a database
-**Administrative** -- who made it and who can use it
- Descriptive and structural meta-data is essential if you want someone else to understand your dataset 
- Descriptive meta-data or "data dictionary" files for raw data are especially useful as a guide to downstream data-tidying
- Check out DataONE's best practicies page for more information:  https://www.dataone.org/best-practices


### 2. Keep a tidy file structure  

- 3 main folders: data, code, output  
- R-synthesized data folder, sometimes  
- RMarkdown file or at least a readme.txt file that describes in human language what your scripts do  
- RProjects can be useful for monitoring files in your project folder  

```{r tidyFolders, echo = TRUE}

# read-in data from the data folder
dat<-read.table("data/metaDataset.txt", header=TRUE, sep="\t")

# make a plot with that data
dat.sub<-subset(dat, measCat=="soiln" & traitCat=="percN" & !is.na(InvSpInvArea_cwm) & !is.na(yi)) #make sure the x and y data point are all there

p<-ggplot(dat.sub, aes(x=InvSpInvArea_cwm, y=yi))  + 
              geom_abline(intercept = 0, slope=0, lty=2) +
              geom_point(shape=19) + 
              xlab("Invasive species Leaf %N") + ylab("Soil N effect size")
p
# save the plot to the output folder
pdf(file = "output/myFigure.pdf", width = 2.5, height = 2.5)
p
dev.off()

```


### 3. Annotate your code

- Hashtag comments within your R scripts
- RMarkdown for more extended descriptions, especially useful for linking results and analyses through text


### 4. Tidy data guidelines

- Don't put multiple variables in one column
- Variables shouldn't be store in both rows and columns
- A table should only have 1 type of observational unit
- R packages that can help: dplyr, tidyr, data.table, reshape2


### 5. Scripting tips

- Start with pseudo code
- Break problems up into small pieces
- Minimize redundancy by using functions and loops
- Set up your workspace the start of the script file with info like the packages required, custom functions, file path identifiers, etc.

```{r scripting, echo = TRUE}

# source in my ggplot theme
source('CODE/mytheme.R')

#make the previous plot pretty and save it as something new
p + mytheme
pdf(file = "output/myFigure_pretty.pdf", width = 2.5, height = 2.5)
p + mytheme
dev.off()

```


